Demystifying the Computer Science \\ PhD Admission in the US\\{\large A Handbook for International \emph{and Domestic} Students

### 2024-03-29 demystify.tex
updates

\subsection{Who is this document for?}

While this document can be useful for all students interested in CS PhD in the US, it is primarily intended for international students from \emph{smaller countries and less well-known universities}. These students have the disadvantage of not having access to the same resources and information as students from larger countries and well-known universities. This document aims to level the playing field by providing information that is not readily available to these students.

@@

While many find CSRankings useful, it is not perfect, and in general, rankings are superficial, and you need to do more research to be informed and make better decisions. For example, you might prefer places with a large community from your country.  You might need to think about the cost of living, e.g., living in California is more expensive than in Virginia.  You might prefer places with high tech companies (Seattle or Silicon Valley), with many outdoor activities (e.g., hiking, skiing), or with better weather (\emph{``PhD can be depressing, so would you rather be depressed in California or in New York?''}).

@@

\paragraph{You aim too high}  This is perhaps the most common one: you have applied to schools that are too high for your profile. While being ambitious is good, you should understand how PhD admission works (e.g., by reading this handbook! and realize things such as having a good GPA doesn't mean much) and be realistic!

You should apply to a range of schools, including ``safety'' ones.   Before applying, you should talk to your professors and ask them to give you an honest opinion on where you should apply.

@@

A related reason is that you did not show enough interests in the program.
For example, your SOP did not mention any faculty or research area at the program, or you did not respond to professors interested in interviewing you (which might be considered unprofessional and burn bridges).

>>>>

\section{Who is this document for?}

While this document can be useful for all students interested in CS PhD in the US, it is primarily intended for \text bf{international students} from \textbf{smaller countries and less well-known universities} (similarly, for domestic students from universities with no PhD programs or limited research experiences).  
This document aims to level the playing field by providing information that is not readily available to these students.

@@

While many find CSRankings useful, it is not perfect, and in general, rankings are superficial, and you need to do more research to be informed and make better decisions. For example, you might prefer places with a large community from your country (e.g., Northern VA has a large population of Vietnamese). You might prefer places with high-tech companies (Seattle or Silicon Valley), with many outdoor activities (e.g., hiking, skiing), or with better weather (\emph{``PhD can be depressing, so would you rather be depressed in California or New York?''}).  You might need to think about the cost of living, e.g., living in California is more expensive than in Virginia, or places with high crime rates (note that while some university might be in a high crime city, but the campus itself is safe).

@@

\paragraph{You aim too high}  This is a very common reason: you have applied to schools that are too high for your profile. While being ambitious is good, you should understand how PhD admission works (e.g., by reading this handbook and realizing things such as having a good GPA doesn't mean much) and be realistic!

Before applying, you should talk to your professors and ask them to give you an honest opinion on where you should apply. You should also apply to a range of schools, including ``safety'' ones.

@@

A related reason is that you did not show enough interests in the program.
For example, your SOP did not mention any faculty or research area at the program, you did not respond to professors interested in interviewing you (which might be considered unprofessional and burn bridges), or during the interview you show little interest or enthusiasm (because they are your safety schools) in the program or profs.
